---
title: "Session 2 - Sampling I: SRS, Stratified, and Cluster Sampling"
author: "ISM2324"
output:
  pdf_document: default
  html_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goals
1. Learn to know the population frame, target variable, and context of the Swiss minaret initiative
2. Draw simple random samples, stratified samples, and cluster samples using the `sampling` package (Tillé and Matei 2016)
3. Estimate quantities of interest and their sampling variance using the `survey` package (Lumley 2010)


## Preparations

### Clean up your global environment
```{r}
rm(list=ls())
```


### Install and load packages
* `pacman` for package management
```{r, message=FALSE}
#  install.packages("pacman")
```

* Uncomment if you need to install `pacman`
* `tidyverse`: package for all tidy operations (dplyr, ggplot2, purrr, ...)
* `sampling`: Tillé and Matei's package for survey sampling
* `survey`: Lumley's package for analyzing complex surveys

```{r, message=FALSE}
pacman::p_load(tidyverse,sampling,survey)
```

### Set working directory
* Remember to use slashes and not backslashes!

```{r, message=FALSE}
#setwd('C:/Users/Peter/Nextcloud/LEHRE/ISM2324/Tutorial/')
```

### Load and inspect data
We are using (synthetic) population data from an initiative on banning the construction of minarets which was held in Switzerland on 2009/11/29 (see https://swissvotes.ch/vote/547.00 for more information). The data set has one row for each of the ~2.7M voters. The variable of interest is `vote` (1 yes, 0 no), the stratum variable is `canton_id`, the cluster variable is `mun_id`. Our aim is to estimate the total number and proportion of yes-votes based on different sampling schemes.   

```{r, message=FALSE}
load("minaret.RDa")
head(minaret)
summary(minaret)
```

<mark>Your task: Inspect the data, retrieve the number of voters, clusters, and strata 

```{r, message=FALSE}
dim(minaret)
length(unique(minaret$canton_id))
length(unique(minaret$mun_id))
```

* Add `voter_id`, number of cantons and municipality, population sizes overall, per canton, and per municipality for later use
```{r, message=FALSE}
minaret <- minaret %>% 
  mutate(voter_id = seq(1, n(), by = 1), N = n(), H = n_distinct(canton_id), M = n_distinct(mun_id)) %>%
  group_by(canton_id) %>%  mutate(N_h = n()) %>% ungroup() %>%
  group_by(mun_id) %>%  mutate(N_k = n()) %>% ungroup()
```

## The sampling package (Tillé and Matei 2016)

```{r, message=FALSE}
ls("package:sampling")
```
### The most important functions for this session:
* `srswor()` draws a simple random sampling without replacement of size n
* `inclusionprobabilities()` computes the first-order inclusion probabilities
* `inclusionprobastrata()` computes inclusion probabilities for stratified samples
* `cluster()` draws cluster samples with equal/unequal probabilities
* `strata()` draws stratified samples with equal/unequal probabilities
* `mstage()` implements multistage sampling with equal/unequal probabilities
* `getdata()` extracts the observed data from a data frame. The function is used after a sample has been drawn


## The survey package (Lumley 2010)
```{r, message=FALSE}
ls("package:survey")
```
### The most important functions:
* `svydesign` assigns sampling design information to a data frame containing a sample 
* There are many associated functions (`svytotal()', svymean()`, `svyvar()`,`svyquantile()` among many others)

## Simple random sampling 

* First, fix the starting value to make random number generation replicable.
```{r, message=FALSE}
set.seed(123)
```

* Overall population size
```{r, message=FALSE}
N <- nrow(minaret)
```

* Overall sample size
```{r, message=FALSE}
n <- 1000
```

* Draw a SRS of size $n=1000$ without replacement from our minaret frame.
```{r, message=FALSE}
srs_sample <- srswor(n = n, N = N)
srs_sample <- getdata(data = minaret, m = srs_sample)
```

* Specify a `svydesign` object.
```{r, message=FALSE}
srs_design <- svydesign(ids = ~1, data = srs_sample,
                        fpc = ~N)
summary(srs_design)
```

* Estimate population total of voters who voted in favor of banning the construction of minarets; compare to actual total
```{r, message=FALSE}
svytotal(x = ~vote, design = srs_design)

sum(minaret$vote)
```
* Confidence interval (default is 95%)
```{r, message=FALSE}
confint(svytotal(x = ~vote, design = srs_design))
```

* Estimate population proportion of voters who voted in favor of banning the construction of minarets; compare to actual proportion
```{r, message=FALSE}
svymean(x = ~vote, design = srs_design)

sum(minaret$vote)/length(minaret$vote)
```
* Normal approximation to confidence interval as above or `svyciprop` specifically for proportions near 0 or 1. Function uses different methods, see help file. 
```{r, message=FALSE}
confint(svymean(x = ~vote, design = srs_design))
svyciprop(formula = ~I(vote==1), design = srs_design)
```



## Stratified random sampling

<mark>Your task: Use the `group_by()` and `summarise()` functions to display the proportions of Yes votes by canton

```{r, message=FALSE}
vote_by_canton <- minaret %>% 
  group_by(canton_id) %>% 
  summarise(p_yes=mean(vote),
          v_yes=var(vote))
vote_by_canton
```


* Define the number of strata $H=26$
```{r, message=FALSE}
H <- length(unique(minaret$canton_id))
```

* Draw a stratified sample of size $n = 1000$, equal sample size $n/H$ in each stratum. Attention: the function name `strata` is ambiguous. We therefore need to specify the package.  
```{r, message=FALSE}
str_sample_eqn <- sampling::strata(data = minaret, stratanames = "canton_id", 
                     size = round(rep(n/H, times = H)), method = "srswor")
str_sample_eqn <- getdata(data = minaret, m = str_sample_eqn)
``` 

* Specify `svydesign` object
```{r, message=FALSE}
str_design_eqn <- svydesign(ids = ~1, strata = ~canton_id, fpc = ~N_h, data = str_sample_eqn) 
```

* Estimate population share of yes votes, compare to population parameter
```{r, message=FALSE}
svymean(~vote, design = str_design_eqn)

sum(minaret$vote)/length(minaret$vote)
```

* Calculate design effect
```{r, message=FALSE}
svymean(~vote, design = str_design_eqn, deff = TRUE)
```

* Draw a stratified sample of size $n = 1000$ with sample size allocation proportionate to stratum size $n_h = n \frac{N_h}{N}$. We use the `inclusionprobastrata` function which (somewhat redundantly?) requires $n_h$. The method used is systematic sampling. Recall the Kalton (1983) example from the lecture.

* Stratum size in the population
```{r, message=FALSE}
N_h <- table(minaret$canton_id)
```


* Population fraction in stratum
```{r, message=FALSE}
p_h <- N_h/unique(minaret$N)

pi_str_pps <- inclusionprobastrata(minaret$canton_id,n*p_h)

str_sample_pps <- sampling::strata(data = minaret, stratanames = "canton_id", 
                     size = n*p_h, pik = pi_str_pps, method = "systematic")
str_sample_pps <- getdata(data = minaret, m = str_sample_pps)
``` 

* Specify `svydesign` object
```{r, message=FALSE}
str_design_pps <- svydesign(ids = ~1, strata = ~canton_id, fpc = ~N_h, data = str_sample_pps) 
```

* Estimate population share of yes votes, compare to population parameter, 
```{r, message=FALSE}
svymean(~vote, design = str_design_pps, deff = TRUE)

sum(minaret$vote)/length(minaret$vote)
```


## Single-stage cluster sampling

* Our clusters $k$ are the Swiss municipalities.

* Define the number of clusters $M=2144$
```{r, message=FALSE}
M <- length(unique(minaret$mun_id))
```

* Number of clusters to be sampled
```{r, message=FALSE}
m <- 50
```

* Draw cluster sample SRSWOR
```{r, message=FALSE}
cls_sample <- sampling::cluster(data=minaret, clustername="mun_id", size=m, method="srswor")
cls_sample <- getdata(minaret, cls_sample)
nrow(cls_sample)
```
* Define survey design object
```{r, message=FALSE}
cls_design <- svydesign(ids = ~mun_id, data = cls_sample,
                        fpc = ~M)
summary(cls_design)
```

* Estimate population share of yes votes, design effect
```{r, message=FALSE}
svymean(~vote, design = cls_design, deff = TRUE)
```

<mark>Your task: Compute and interpret the effective sample size (you can use the `deff()`function from the `survey` package to extract the value of the design effect)
```{r, message=FALSE}
phat_cls <- svymean(~vote, design = cls_design, deff = TRUE)
nrow(cls_sample)
deff(phat_cls)
nrow(cls_sample)/deff(phat_cls)
```


# This is already for next week!

## Two-stage cluster sampling

* Cluster population sizes
```{r, message=FALSE}
N_k <- table(minaret$mun_id)
str(N_k)
```

* Stage 1 inclusion probabilities proportionate to size (PPS)
```{r, message=FALSE}
pi1_tss <- inclusionprobabilities(N_k, m)
```

* Draw the sample. Note that the second stage is specified as stratified sampling (among the clusters selected at stage 1). Sample sizes are constant at stage 2, $n_k=n/m$.
```{r, message=FALSE}
tss_sample <- mstage(data = minaret, stage = list("cluster", "stratified"), 
                     varnames = list("mun_id", "voter_id"), size = list(size1 = m, size2 = c(round(rep(n/m, m)))), 
                     method=list("systematic","srswor"), pik = list(pi1_tss))
tss_sample <- getdata(data = minaret, m = tss_sample)[[2]]
```

* Append inclusion probabilities to sample data frame
```{r, message=FALSE}
tss_sample <- tss_sample %>%
  mutate(pi1 = inclusionprobabilities(N_k, m), pi2 = n/m/tss_sample$N_k) 
```

* Inspect the sample
```{r, message=FALSE}
View(tss_sample)
```

* Specify a survey design object, Brewer is a method to approximate (under the assumption of sampling with replacement) second-order or joint inclusion probabilities for elements i and j. This is necessary to calculate the sampling variance of the estimator.
```{r, message=FALSE}
tss_design <- svydesign(ids = ~mun_id+voter_id, probs = ~pi1+pi2, pps = 'brewer', data = tss_sample)
summary(tss_design)
```

* Estimate population share of yes votes, compare to population parameter, 
```{r, message=FALSE}
svymean(~vote, design = tss_design, deff = TRUE)
sum(minaret$vote)/length(minaret$vote)
```