---
title: "Session 4 - Nonresponse adjustments: imputation methods"
author: "ISM2425"
output:
  html_document: default
  pdf_document: default
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Goals
* Learn how to implement imputation methods using the exit poll data from Selb, Goebel and Lachat (2021)

## Preparations
We need many of the objects from the previous session. So we run all the R code up to and including line 152 of `Session05_Adjustment.Rmd` in a single chunk, and we prevent RMarkdown from including the output by setting `include` argument to `FALSE`. 


```{r, include=FALSE}
rm(list=ls())

pacman::p_load(tidyverse,survey)

load("FPEsurvey.rda")

cand_names <- c("Arthaud","Poutou","Melenchon","Hamon","Macron","Fillon","Dupont-Aignon","Le Pen","Cheminade","Asselineau","Lassalle")

age_groups <- c("18-24","25-34","35-44","45-54","55-64","65-74",">75")

resp_data <- resp_data %>% 
  mutate(choice = factor(choice, levels = seq(1,11,1), labels = cand_names), age = factor(age, levels = seq(1,7,1), labels = age_groups), sex = recode_factor(sex, `1` = "female", `2` = "male")) %>%
  mutate_at(vars(Arthaud:Lassalle), ~na_if(., 0)) %>%
  filter(!is.na(choice))

multi_design <- svydesign(id = ~id + resp_id, probs = ~first_stage + second_stage, data = resp_data)

cand_totals <- data.frame(choice = cand_names, totals = c(232384,394505,7059951,2291288,8656346,7212995,1695000,7678491,65586,332547,435301))

post_design <- postStratify(design = multi_design, strata = ~choice, population = cand_totals)   
resp_data <- resp_data %>%
  mutate(
    runoff = case_when(
      choice == "Macron" ~ "Macron",
      choice == "Le Pen" ~ "Le Pen",
      Macron > LePen ~ "Macron",
      LePen > Macron ~ "Le Pen",
      TRUE ~ NA_character_
    )
  )

multi_design <- svydesign(id = ~id + resp_id, ~first_stage + second_stage, data = resp_data)

post_design <- postStratify(design = multi_design, strata = ~choice, population = cand_totals)
```

## Background 1
Recall that, in our first try to redistribute the votes for eliminated candidates, we make several simplifying assumptions: 

* Round-one votes for Macron and Le Pen would remain stable
* Those who voted for another candidate in round 1 would vote for the higher-rated among the runoff candidates
* Those with tied (or missing) ratings among the runoff candidates will abstain

As it turned out, these were many. Although our prediction of the runoff vote shares based on the poststratified sample was quite accurate, we massively overestimated abstention. More than a quarter of our respondents had missing values on the `runoff`variable and were therefore predicted to abstain: 

```{r, message=FALSE}
sum(is.na(resp_data$runoff))

sum(is.na(resp_data$runoff))/nrow(resp_data)
```

### Install and load packages
```{r, message=FALSE}
pacman::p_load(VIM,mitools,mice)
```

* `VIM`: visualization and imputation of missing data
* `mitools`: handles multiply imputed data sets with survey functions
* `mice`: multiple imputation by chained equations; an iterative regression-based imputation algorithm due to White, Royston and Wood (2011).  
* `survey`: also includes functions for poststratification and raking

## Background 2
Recall from the lecture:

1. Standard regression-based (deterministic) imputation

- Regress incomplete y on complete x
- Impute linear predictor, xbhat
- Problems: prediction uncertainty, incomplete x, imputation uncertainty?

2. Stochastic imputation
- Regress incomplete y on complete x
- Impute linear predictor, xbhat, plus random draw from the residual distribution, N(0,shat2_e)
- Problems: incomplete x, imputation uncertainty?
3. Iterated imputation
- Impute missing x by random draws from N(xbar,shat2_x), stochastically impute y
- Unplug imputed values from x, use imputed y to impute x
- Unplug imputed y, use updated imputed x to impute y etc.
- Problems: imputation uncertainty
4. Multiple imputation
- Repeat iterated imputation m times
- Perform analysis with m data sets
- Average over m trials
- Imputation uncertainty: imputation variance over m trials 

We will only do state-of-the-art multiple imputation using the `mice` package. Single (stochastic) imputation can be implemented with `mice` by setting $m = 1$. 

## Inspect missingness patterns
```{r, message=FALSE}
aggr(resp_data[,c(4:17)], prop = FALSE)   
```
* We used the `aggr()` function from the `VIM`package to plot missing data statistics and patterns
* In doing so, we exclude ID variables and weights in columns 1-3 and 18-20 ; they don't have missing values


## Covariate selection
The `mice` function requires a so-called predictor matrix - a numeric matrix of `length(blocks)` rows and `ncol(data)` columns, containing 0/1 data specifying the set of predictors to be used for each target column. Each row corresponds to a variable block, i.e., a set of variables to be imputed. A value of 1 means that the column variable is used as a predictor for the target block (in the rows). Here we only use the candidate ratings in columns 4-14 of `resp_data`, as they can be reasonably modelled using linear regression. We could also specify other models for categorical outcomes (such as choice, sex, age group, or polling place). 

```{r, message=FALSE}
predictor_matrix <- matrix(data = c(0,0,0,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0), nrow = ncol(resp_data), ncol = ncol(resp_data), dimnames = list(colnames(resp_data), colnames(resp_data)),  byrow = TRUE)

predictor_matrix
```
## Imputation of candidate ratings using linear regression (method "norm")
```{r, message=FALSE}
imps <- mice(resp_data, m=5, method = c("", "", "", "norm", "norm", "norm", "norm", "norm", "norm", "norm", "norm", "norm", "norm", "norm", "", "", "", "", "", "", ""), predictorMatrix = predictor_matrix, seed=33)
```

* Put the imputations and the original data together in a list of m data frames (we need to use base R functions for data management here)
```{r, message=FALSE}
imp_data <- imputationList(lapply(seq(imps$m), complete, data=imps))
```

## Re-calculate the `runoff` variable
I have tried to write this code using `dplyr` functions but apparently, this doesn't work with objects of class `imputationList`. 

```{r, message=FALSE}
for (i in seq(imps$m)) {
  imp_data$imputations[[i]]$runoff <- ifelse  (imp_data$imputations[[i]]$choice=="Macron", 1, ifelse (imp_data$imputations[[i]]$choice=="Le Pen", 2, NA))
  
  imp_data$imputations[[i]]$runoff[!is.na(imp_data$imputations[[i]]$Macron)&!is.na(imp_data$imputations[[i]]$LePen)&imp_data$imputations[[i]]$Macron > imp_data$imputations[[i]]$LePen] <- 1
  
  imp_data$imputations[[i]]$runoff[!is.na(imp_data$imputations[[i]]$Macron)&!is.na(imp_data$imputations[[i]]$LePen)&imp_data$imputations[[i]]$Macron < imp_data$imputations[[i]]$LePen] <- 2
  
  imp_data$imputations[[i]]$runoff <- factor(imp_data$imputations[[i]]$runoff, levels = c(1,2), labels = c("Macron","Le Pen"))
}
```
This code iterates through each imputed dataset indexed by `i` in the sequence of `imps$m`.
1. For first-round voters of Macron or Le Pen, it assigns numeric values to the runoff variable based on the choice variable. If choice is "Macron", runoff is set to 1; if choice is "Le Pen", runoff is set to 2; otherwise, it's set to NA.
2. For voters of eliminated candidates, it refines the runoff variable based on conditions related to Macron and LePen variables within each imputed dataset. If Macron and LePen are not missing and Macron is greater than LePen, runoff is set to 1. If Macron and LePen are not missing and Macron is less than LePen, runoff is set to 2.
3. The numeric runoff values are converted back to a factor variable where 1 corresponds to "Macron" and 2 corresponds to "Le Pen". This step sets the levels and labels accordingly.

## Specify sampling design object
```{r, message=FALSE}
imp_designs <- svydesign(id = ~id + resp_id, ~first_stage + second_stage, data = imp_data)
```

## Estimate / predict the population totals and vote shares for Macron in the runoff
```{r, message=FALSE}
imp_shares <- with(imp_designs, svymean(~runoff, na.rm = TRUE))
imp_totals <- with(imp_designs, svytotal(~runoff, na.rm = TRUE)) 

summary(MIcombine(imp_shares))
summary(MIcombine(imp_totals))
```

# Poststratification
Note: `postStratify()` does not take a list object generated by `imputationList()` as an argument, therefore we need to specify the sampling design separately for each imputed data set

```{r, message=FALSE}
imp_post_designs <- list()
imp_post_shares <- list()
imp_post_totals <- list()

for (i in seq(imps$m)) {
  imp_post_designs[[i]] <- postStratify(imp_designs$designs[[i]], strata = ~choice, population = cand_totals)
  imp_post_shares[[i]] <- svymean(~runoff, imp_post_designs[[i]], na.rm = TRUE)
  imp_post_totals[[i]] <- svytotal(~runoff, imp_post_designs[[i]], na.rm = TRUE)

}

summary(MIcombine(imp_post_shares))
summary(MIcombine(imp_post_totals))
```


* Actual runoff result: Macron 0.661, Le Pen 0.339. Not bad. In the paper (Selb et al. 2021) we used polytomous instead of linear regression, which further improved the prediction somewhat.  


## Hotdeck imputation
Finally, let's try out the `hotdeck()` function rom the `VIM` package (Kowarik and Temple 2016)

```{r, message=FALSE}
hot_data <- hotdeck(data = resp_data, variable = c("Macron", "LePen"), imp_var = TRUE)

View(hot_data)
``` 

Now the redistribution step:
```{r, message=FALSE}
hot_data <- hot_data %>%
  mutate(
    runoff = case_when(
      choice == "Macron" ~ "Macron",
      choice == "Le Pen" ~ "Le Pen",
      Macron > LePen ~ "Macron",
      LePen > Macron ~ "Le Pen",
      TRUE ~ NA_character_
    )
  )
```

Specify the design object, estimate 
```{r, message=FALSE}
hot_design <- svydesign(id = ~id + resp_id, ~first_stage + second_stage, data = hot_data)

post_hot_design <- postStratify(design = hot_design, strata = ~choice, population = cand_totals)

svymean(~runoff, hot_design, na.rm=T)
svytotal(~runoff, hot_design, na.rm=T)
svymean(~runoff, post_hot_design, na.rm=T)
svytotal(~runoff, post_hot_design, na.rm=T)
```

## Final thoughts
Note that we use observed relationships between (or, in the hotdeck case, patterns across) variables (candidate ratings, choice, age groups, gender, polling place) to impute missing Macron and Le Pen ratings. In a way, we let those who haven't made up their minds about Macron and Le Pen learn what other, similar voters already knew at the first election round.  